---
title: "Exercise 7.1 Linear Regression"
author: "James Hooi"
date: "October 25, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
states.data <- readRDS("dataSets/states.rds")
```

## Least squares regression

Use the *states.rds* data set. Fit a model predicting energy consumed per capita (energy) from the percentage of residents living in metropolitan areas (metro).

Check out the base data and plot energy vs metro:

```{r plot1, echo=TRUE, eval=TRUE, warning=FALSE}
plot(energy ~ metro, data = states.data)
```

Looks like a slightly negative relationship, does a linear regression support this?

```{r mdl1, echo=TRUE, eval=TRUE, warning=FALSE}
energyReg <- lm(energy ~ metro, data = states.data)
summary(energyReg)
```

The least squares regression supports metro having a slight negative relationship with energy. metro is a significant variable with respect to energy, however the R^2^ and adjusted R^2^ are both quite low.

```{r mdleval1, echo=TRUE, eval=TRUE, warning=FALSE}
par(mar = c(4, 4, 2, 2), mfrow = c(1, 2))
plot(energyReg, which = c(1, 2))
```

The model becomes a less reliable predictor at higher energy levels, i.e. towards the left of the raw data plot energy v metro. This is expected as the data is sparse here.

### Try adding additional predictor

I added density to the model to see if it improves or degrades.

```{r plot2, echo=TRUE, eval=TRUE, warning=FALSE}
par(mar = c(4, 4, 2, 2), mfrow = c(1, 2))
plot(energy ~ metro, data = states.data)
plot(energy ~ density, data = states.data)
```

Both look like a slightly negative relationship, although the data is clustered around different areas.

```{r mdl2, echo=TRUE, eval=TRUE, warning=FALSE}
energyReg2 <- lm(energy ~ metro + density, data = states.data)
summary(energyReg2)
```

The least squares regression supports both variables having a negative relationship with energy, however neither is significant in spite of the R^2^ and adjusted R^2^ improving.

```{r mdleval2, echo=TRUE, eval=TRUE, warning=FALSE}
par(mar = c(4, 4, 2, 2), mfrow = c(1, 2))
plot(energyReg2, which = c(1, 2))
```

As before, the model becomes a less reliable predictor at higher energy levels, i.e. towards the left of the raw data plot energy v metro + density.


## Interactions

Tested the interaction between metro and density variables.

```{r inter1, echo=TRUE, eval=TRUE, warning=FALSE}
energyReg3 <- lm(energy ~ metro*density, data = states.data)
coef(summary(energyReg3))
```

The slope between the two variables is close to zero so they have a minimal impact on each other.

Looking at regions, it does appear there are significant differences in the relationship between energy and metro given each region. While all have a negative relationship, the slopes of the lines are quite different (as shown in the following plot).

```{r inter2, echo=TRUE, eval=TRUE, warning=FALSE}
coef(summary(lm(energy ~ metro + region, data = states.data)))
ggplot(states.data[which(!is.na(states.data$region)),], aes(x=metro, y=energy, col=region)) +
  geom_point() + facet_grid(. ~ region) + stat_smooth(method = "lm", se = FALSE, col="black")
```

Seeing if region improves the model:

```{r final, echo=TRUE, eval=TRUE, warning=FALSE}
energyReg4 <- lm(energy ~ metro + region, data = states.data)
summary(energyReg4)
```

The R^2^ and adjusted R^2^ both go up although two of the regions are not significant predictors. Adding other variables does not improve on this model.